# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data  
Here's the R-code for loading and preprocessing the data:  
```{r LoadnPreprocess}
act <- read.csv("activity.csv")
actint <- act$interval
actdate <- act$date
actsteps <- as.integer(act$steps)
library(data.table)
actt <- data.table(act)
```

## What is mean total number of steps taken per day?  
Here's the R-code for computing and plotting the mean total number of steps taken per day:  
```{r TotalStepsPerDay }
sumv <- actt[,as.double(sum(steps,na.rm = TRUE)),by = date]
sumv1 <- sumv[[1]]
sumv2 <- sumv[[2]]
meansum <- mean(sumv2)
mediansum <-  median(sumv2)
hist(sumv2, breaks =length(sumv2), col = "red", main = "Fig. 1. Total steps per day",xlab ="Total steps per day")
```
### The *mean* and *median* of total number of steps taken per day are *`r meansum`* and *`r mediansum`* respectively.

## What is the average daily activity pattern?  
Here's the R-code for computing and plotting the mean total number of steps taken per 5-min interval across all days:  
```{r AverageStepsPer5minInterval}

avgint <- actt[,as.double(mean(steps,na.rm = TRUE)),by = interval]
avgint1 <- avgint[[1]]
avgint2 <- avgint[[2]]
plot(avgint1,avgint2, type ="l",main = "Fig. 2. Average Steps per 5-min Interval",xlab = "Intervals",ylab = "Average Number of Steps")
maxstepsint <- avgint1[which.max(avgint2)]
```
### On average across all the days in the dataset, *the 5-minute interval = `r maxstepsint`* contains the maximum number of steps.

## Inputing missing values  
Here's the R-code for replacing missing data with the mean for the corresponding 5-minute interval:    
```{r}
actna <- as.numeric(is.na(actsteps))
avgint2r <- rep(avgint2, times=length(sumv2))
actstepsn <- rowSums(cbind(actsteps,actna*avgint2r),na.rm = TRUE)
missingvalrowcnt <- sum(actna)
```
### The total number of missing values in the dataset (i.e. the total number of rows with NAs) is `r missingvalrowcnt`.

## What is new  mean total number of steps taken per day?
Here's the R-code for computing and plotting the New mean total number of steps taken per day:
```{r TotalStepsPerDayNew}
actnewt <- data.table(date=actdate,interval = actint, steps = actstepsn)

sumnewv <- actnewt[,as.double(sum(steps,na.rm = TRUE)),by = date]
sumnewv1 <- sumnewv[[1]]
sumnewv2 <- sumnewv[[2]]
meansumnew <- mean(sumnewv2)
mediansumnew <-  median(sumnewv2)
hist(sumnewv2, breaks =length(sumnewv2), col = "red", main = "Fig. 3. Total steps per day (after replacing NA values)",xlab ="Total steps per day")
```
### The NEW *mean* and *median* of total number of steps taken per day are *`r meansumnew`* and *`r mediansumnew`* respectively.

### The impact of replacing missing data with the mean for the corresponding 5-minute interval, on the estimates of the total daily number of steps is as follows:
**1.  The new Mean is equal to the new Median of the total number of steps taken per day whereas the old Mean is NOT equal to the old Median.**  
**2.	The new Mean and Median are greater than the old ones.**

## Are there differences in activity patterns between weekdays and weekends?  
Here's the R-code for computing (using the NEW data) and plotting the number of steps taken per 5-min interval, averaged across all weekday days or weekend days:  
```{r weekdayweekendpanel}
wday <- weekdays(strptime(actdate, "%Y-%m-%d"))
wd <- rep ("weekday", times = length(avgint1))
we <- rep ("weekend", times = length(avgint1))
wdayint <- c(wd,we)
actinttmsw <- c(avgint1, avgint1)

wdc <- as.numeric(!(wday == "Saturday") & !(wday == "Sunday"))
wec <- as.numeric((wday == "Saturday") | (wday == "Sunday"))
actnewtwd <- data.table(date=actdate,interval = actint, steps = actstepsn*wdc)
actnewtwe <- data.table(date=actdate,interval = actint, steps = actstepsn*wec)
avgintwd <- actnewtwd[,as.double(mean(steps,na.rm = TRUE)),by = interval]
avgintwe <- actnewtwe[,as.double(mean(steps,na.rm = TRUE)),by = interval]
avgintn <- c(avgintwd[[2]],avgintwe[[2]])
maxstepsintn <- actinttmsw[which.max(avgintn)]

weekint <- data.frame(interval = actinttmsw,f = factor(wdayint),avgSteps = avgintn)
library(lattice)
p <- xyplot(avgSteps~interval| f, data = weekint, type = "l", layout = c(1,2), scales=list(x=list(tick.number = 10)),xlab = "Intervals",ylab="Average Number of Steps",main = "Fig. 4. Average Steps per 5-min Interval") ## plot w 2 panels
print(p)

```
### On average across all the days in the NEW dataset, *the 5-minute interval = `r maxstepsintn`* contains the maximum number of steps.  
### The average number of steps per 5-minute interval on weekdays seem to be greater than those on weekends.
The End